Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2020-05-27 Manuscript/manuscript.Rmd
added to discussion

???stopped here

# Conclusion

In summary, we have shown that TAD boundary prediction is a multi-faceted problem composed of many data irregularities and characteristics that previous methods have failed to address. Through our ensemble framework, we provide a useful guide toward the exploration of chromatin organization, while highlighting the impact of data resolution, class imbalance, and feature engineering in TAD boundary prediction. Lastly, by building models using known enriched transcription factors, we were able to bridge the resolution gap between 1D ChIP-Seq annotations and 3D Hi-C sequencing for more precise TAD boundary prediction. This improved precision in TAD boundary location offers the potential for further exploration into the 3D organization of the human genome.

>>>>

In summary, we have shown that TAD boundary prediction is a multi-faceted problem composed of many data irregularities and characteristics. Simply quering a contact matrix ignores the fundamental roles of known molecular drivers of 3D chromatin. Instead, we propose a method for turning TAD-calling into a supervised learning framework built using ChIP-seq annotations. Specifically, we introduce two novel concepts known as *shifted binning* and *distance*-type predictors which we use to build random forest classification models for predicting boundary regions. We then introduce *preciseTAD*, a method for leveraging random forests to predict TAD(loop) boundaries at base pair resolution. Our method is able to bridge the resolution gap between 1D ChIP-seq annotations and 3D Hi-C sequencing data for more precise and biologically meaningful boundary prediction. By releasing our method as an open source R package, we hope to provide an efficient and easy-to-use tool to further explore the 3D organization of the human genome.
